---
title: "Database Connections in R"
author: "Chris Mainey chris.mainey@uhb.nhs.uk"
date: "2018/01/28 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    css: "libs/CM_CSS.css"
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.align = "center", 
                      dev.args = list(png = list(type = "cairo")), fig.retina=3) 

library(Cairo)
library(RODBC)
#Connection string
RODBC_connection <- odbcDriverConnect(paste('driver={SQL Server};server=',
                                            Sys.getenv("SERVER"),
                                            ';database=',
                                            Sys.getenv("DATABASE"),
                                            ';trusted_connection=true', sep = ""))

```



<br><br><br><br><br><br>
# Database connections in R

<br>

__Chris Mainey__ 

<a href="mailto:chris.mainey@uhb.nhs.uk"> chris.mainey@uhb.nhs.uk</a>

<br><br>
2019/06/20 (updated: `r Sys.Date()`)
<br><br><br><br><br><br>
<span style="font-size:14px !important;"> __&copy; Healthcare Evaluation Data (HED)__- part of Health Informatics, University Hospitals Birmingham NHS Foundation Trust.  
<b>NOT TO BE REPUBLISHED OR DISTRIBUTED WITHOUT CONSENT</b> </span>


---
  
# Two common methods
  
There are two common methods of connection, both of which use Open Database Connectivity (ODBC) drivers:
  
1.  The `RODBC` package.
2.  The `DBI` system, using `dplyr,` `dbplyr` and `odbc`.

<br>
+ Both of these create a connection, using a 'connection string' 
+ This can be used to create a connection object
+ We can use the object to pull data into `R` or manipulate it in the database.

---
  
## 1.  `RODBC`
  
+ This is the simpler of the two interfaces, and uses slightly older code.
+ It can be used to connect to anything that uses Open Database Connectivity (ODBC).  
<br>
  
```{r ROBCIMport1, eval=FALSE, collapse=TRUE, include=FALSE}
library("RODBC")

#Connection string
# e.g. with a server called "Cliff" and a database called "Richard" your string would be:

RODBC_connection <- odbcDriverConnect('driver={SQL Server};server=Cliff;database=Richard;trusted_connection=true')

dt1 <- sqlFetch(channel=RODBC_connection, sqtable = "MyTable")

```

```{r RODBCImport2, collapse=TRUE}
# Load data from SQL query
dt2 <- sqlQuery(channel=RODBC_connection, query = "select TOP 10 * from MyTable")

dt2

```

---
  
# What is going on here? 
  
+ `trusted_connection=true` passes your windows credentials to the server
+ You can, instead, specify a username (`uid`) and a password (`pwd`)

<br><br>
+ You can also use `RODBC` to write back to database tables, choosing to append or not:
  
```{r RODBCExport1, eval=FALSE}

sqlSave(channel = RODBC_connection, dat = dt2, tablename = "Mytable_version2"
        , append = FALSE, safer = FALSE)

```

---
  
# Other functions
  
There are lots of other functions included with `RODBC` to allow you to see structures etc.  The package vignette is a very helpful place to go for this, along with the help files.

<br><br>
Remember to disconnect at the end of your session:
  
```{r RODBCDisconnect, eval=FALSE}
odbcClose(RODBC_connection)
```


---
  
<br>
  
## 2. `DBI` \ `dplyr`
  
+ `DBI` implements a common database interface in `R`.
+ Can be used with different 'backend' drivers such as MySQL, SQL Server, SQLite, Oracle etc.
+ Works with `tidyverse` functions, including `dplyr` verbs and the pipe `%>%`
+ Faster than `RODBC` to import data
+ Can be used to work with data in the database, without importing it into `R`.
+ `DBI` can be used on it's own, but can be combined with `dplyr`
<br><br>


---

# DBI connection

Requires a different connection string and a few more packages to use:
+ `DBI` - a common Database Interface engine for use in `S` and `R` (see <a href="https://cran.r-project.org/web/packages/DBI/vignettes/DBI-1.html">here</a>)
+ `dplyr` - to make the `tbl` and use it, we'll work with `dplyr` syntax.
+ `dbplyr` - this add-on package allows translation from `dplyr` to SQL.
+ `odbc`- provides the odbc drivers, but you could use the functions below with other drivers instead.

```{r DBI1, message=FALSE, warning=FALSE}
library(dplyr)
library(dbplyr)
library(odbc)
library(DBI)

DBI_Connection <- dbConnect(odbc(), 
                            driver = "SQL Server",
                            server=Sys.getenv("SERVER"),
                            database=Sys.getenv("DATABASE")
)

```

---
  
# Using tables in the database
  
Now we can define a table as if it was part of our `R` workspace, using the connection object and the names of the table in the database. 
+ Do this with `tbl`
+ `glimpse` is a useful function that shows you a summary
<br><br>
```{r dplyr4, message=FALSE, warning=FALSE, collapse=TRUE}
MyTable<-tbl(DBI_Connection, "MyTable")

glimpse(MyTable)

```


---
  
# Constructing `dplyr` query
  
+ We can then perform select or aggregate queries without translation.
+ Even though it returns results, the data are still in the database


```{r dplyr5, message=FALSE, warning=FALSE, collapse=TRUE}

MyTable %>%
  filter(year ==2015) %>%
  group_by(month) %>%
  summarise(AvgEvents = mean(events),
            MaxEvents = max(events),
            N = n()) %>%
  arrange(month)
```


+ `dplyr` can then be used to do fairly complex things in just a few lines.

---
  
# Example: 
+ I'm filtering the data for 2015 and passing it directly into `ggplot2` (tomorrow's session)

```{r DBIplot2, warning=FALSE, message=FALSE, fig.height=4, fig.width=5, fig.retina=2, fig.align = "center" }
library(ggplot2)

MyTable %>%
  filter(year ==2015) %>%
  ggplot(aes(y=events, x=factor(month), group=factor(month))) + 
  geom_boxplot(fill = "dodgerblue2", alpha=0.6, )+
  labs(title = "Monthly Distribution of Events", x="Month", y="Events")


```

---
  
# Using SQL & returning data to `R` (1)
  
+ Can write an SQL query directly using the `dbSendQuery` function.
+ Executes the query on the server-side only.
+ If you want the results back in `R`, you need to use `dbFetch` as well.

```{r dplyrQuery, warning=FALSE, message=FALSE}

SomeRecords <- dbFetch(dbSendQuery(DBI_Connection, "Select TOP 100 * from MyTable"))

#or

SomeRecords <- dbSendQuery(DBI_Connection, "Select TOP 100 * from MyTable") %>%
  dbFetch()

```

---
  
# Using SQL & returning data to `R` (2)
+ May need to pull the data from the server into memory in `R` sometimes.
+ Can do this with `collect`

```{r dplyrCollect, warning=FALSE, message=FALSE}
MyTable_local<- MyTable %>%
  filter(year ==2015) %>%
  group_by(month) %>%
  summarise(AvgEvents = mean(events),
            MaxEvents = max(events),
            N = n()) %>%
  arrange(month) %>%
  collect()

print(MyTable_local)

```

---
  
# Writing to databases
  
You can also write back to a database using the `dbWriteTable` function.
+ For example:
+ Writing a new table current connection, called 'NewDatabaseTable'
+ Using the `R` `data.frame` called "MyTable_local" (that we created in the last section)
+ `append` and `overwrite` options

```{r dplyrWrite, warning=FALSE, message=FALSE, eval=FALSE}

dbWriteTable(DBI_Connection,"NewDatabaseTable", MyTable_local, overwrite=TRUE)

```

---
class: middle, center

# Exercise:
### Follow the instructions in the Excerise file

---

# Useful DBI commands

|Command        | Summary                                                |
| --------------|:------------------------------------------------------:|
|dbConnect()    | Create a DBI connection object                         |
|dbListTables() | List the tables on the connection                      |
|dbListFields() | List the fileds for a given table on a given connection|
|dbSendQuery()  | Send a query to eexecute on the server/connection      |
|dbFetch()      | Fetch the results from the server/connection           |
|dbWriteTable() | Write a table to the connection                        |
|               |                                                        |
|tbl()          | Set a table on the connection as a 'tibble' for `dplyr`|
|glimpse()      | See a summary of the rows, data types and top rows     |

