---
title: "04 Caret Multi-Class Classification"
author: "Gary Hutson - Head of Solutions and AI"
date: "08/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(caretEnsemble)
library(mlbench)
library(DMwR)
library(klaR)
library(magrittr)
library(RSNNS)
library(randomForest)
library(xgboost)
```

## Load data and inspect classification levels

Multi class classification relates to when you dependent variable (predicted variable) has more than two levels. The example we use for this is different readmission bandings. 

```{r load_data, include=TRUE}
df <- read.csv("Data/Multiclass_data.csv", header = TRUE)
head(df, 10)
table(df$Outcome)
df$Outcome <- make.names(df$Outcome)
df$Outcome <- as.factor(df$Outcome)
levels(df$Outcome)

```

Again, this is a test dataset and you might have to undertake some type of over or under sampling to get better representation for each of the class labels. 

## Create test / train split on dataset

I will create a simple partition on the data to examine the class accuracy using a confusion matrix:

```{r split_data, include=TRUE}
split_idx <- caret::createDataPartition(df$Outcome, 
                                        p = 0.8,
                                        list = FALSE)

validation <- df[-split_idx,]
train <- df

```

This uses the concept of a validation dataset to be used after the resampling method to test the accuracy of the model based on new / unseen data. The same concept as we have already examined in the production section of the modelling. 

## Summarise class distribution and dataset

I have created a proportional table function to summarise by the number of classes passed to the table. This will allow you to visualise class membership and proportional distribution:

```{r sum_data_class, include=TRUE}
dim(df)
sapply(df, class)

# Summarise class distribution
class_distribution <- function(field){
  per_dist <- prop.table(table(field)) * 100
  cbind(Class_Frequency=table(field), 
        Class_percentage = per_dist) %>% 
    as.data.frame()
}

class_distribution(df$Outcome)

```

## Set up resampling method

The resampling method to evaluate the model is again 10 fold cross validation. 
```{r resamp_method, include=TRUE}
train_ctrl <- caret::trainControl(method = "cv", number = 10)
metric <- "Accuracy"

```

## Benchmark algorithms
Again, we implement a similar benchmarking procedure as in the regression and classification tasks:

```{r ml_bench, include=TRUE}
set.seed(123)
rf_mod <- caret::train(Outcome ~ ., 
                       data = df, 
                       method = "rf", metric = metric, trainControl=train_ctrl)

# Run on K-Nearest neighbours and naive bayes
set.seed(123)
lda_mod <- caret::train(Outcome ~ ., 
                       data = df, 
                       method = "lda", metric = metric, trainControl=train_ctrl)
set.seed(123)
svm_rad_mod <- caret::train(Outcome ~ ., 
                        data = df, 
                        method = "svmRadial", metric = metric, trainControl=train_ctrl)

```

The algorithms have now run, we will collect the resamples to look at the model fits:

```{r ml_fits, include=TRUE}
mc_results <- caret::resamples(
  list(Random.Forest = rf_mod,
       Linear.Discriminant.Analysis = lda_mod,
       Support.Vector.Machine.Radial.Basis.Kernel = svm_rad_mod)
)

summary(mc_results)
dotplot(mc_results)
print(rf_mod)
```

The accuracy will start to decrease with multiple classification models. For these types of classifications - Deep Neural Networks might be a better option, plus it saves time on exploring features in the models - as these are what I call feature agnostic, as the model finds the relationships. 

## Predict model fits on validation set
With the reserved validation dataset we will estimate the model skill at predicting unseen data:

```{r ml_preds, include=TRUE}
set.seed(123)
mc_pred <- predict(rf_mod, newdata = validation)
head(mc_pred, 10)
```

## Create confusion matrix to evaluate models against validation set

Now, the last step is to observe the confusion matrix fit for this:
```{r con_matrix, include=TRUE}
cm <- caret::confusionMatrix(mc_pred, validation$Outcome)
print(cm)
```

The confusion matrix metrics can then be evaluated to see how well the model performs at predicting associated confusion matrix measures, such as sensitivity (https://classeval.wordpress.com/introduction/basic-evaluation-measures/#:~:text=Sensitivity%20(Recall%20or%20True%20positive,whereas%20the%20worst%20is%200.0.) as well as other associated measures.
